### 2018-04-16 implementation.tex
ok

Address Space Layout Randomization (ASLR) \cite{ASLR}  is probably the most
important source of nondeterminism that arises (only) from change in
process.  ASLR scrambles the layout of

@@

This is a simple, clean example
of a problem that appeared in numerous Python programs when Python
version 3.3 introduced automatic random salting of hashes (see Section
\ref{sec:pnondet} per-process,
in order to mitigate hash-based denial of service attacks
\cite{denial}.  Until version 3.6 this not only resulted in changes in
exact hash values, but in the order of iteration on
dictionaries\footnote{In version 3.6, the default dictionary
  implementation became an ordered dictionary with consistent
  iteration, though actual hashes remained nondeterministic.}.  While
relatively few programs relied on exact hash values, many implicitly
relied on them in that they only functioned correctly with a
predictable order for dictionary iteration across processess.

>>>>

Address Space Layout Randomization (ASLR) \cite{ASLR}  is probably the most
important source of nondeterminism that arises (only) from change in
process.  ASLR scrambles the layout of memory of the process in which
an executable runs in order to make it
harder to exploit memory-safety vulnerabilities.  As a side effect, it
means that a test that, in some circumstances, causes a crash, may at
other times not fail at all.

@@

Another example of process-based nondeterminism was discovered by numerous Python programs when Python
version 3.3 introduced automatic random salting of hashes (see Section
\ref{sec:pnondet} per-process,
in order to mitigate hash-based denial of service attacks
\cite{denial}.  Until version 3.6 this not only resulted in changes in
exact hash values, but in the order of iteration on
dictionaries\footnote{In version 3.6, the default dictionary
  implementation became an ordered dictionary with consistent
  iteration, though actual hashes remained nondeterministic.}.  While
relatively few programs relied on exact hash values, many implicitly
relied on them in that they only functioned correctly with a
predictable order for dictionary iteration across processess.  Testing
Python code for nondeterminism based on this hash seed requires
running in a new process:  Python does not allow changing the salt, since changing hash values on-the-fly would
break all existing dictionaries and other structures relying on hashing.

### 2018-04-16 introduction.tex
ok

This paper proposes (1) a number of formal definitions of types of
nondeterminism and (2) a number of techniques, based on these definitions, for detecting and debugging
nondeterminism in property-based testing, particularly for library
code.  The methods also work for testing at the system level, but are
perhaps most effective for identifying the low-level sources of
nondeterminism in libraries.  This enables a number of useful results:
 the automated production of
nondeterminism-free regression tests even for libraries with some
sources of nondeterminism, the documentation of library
nondeterminism, and, most importantly, detection and effective debugging of
unexpected nondeterminism that is, itself, a bug.  We first
describe the types of nondeterminism we aim to detect, and the
extension of a widely used debugging tool to better handle
nondeterminism as a test property, then describe an implementation of
these ideas in a Python testing tool that is used on real-world
projects.  Finally, we use a set
of case studies of real-world Python libraries to demonstrate the
utility of our ideas and understand the cost of adding nondeterminism
detection to existing automated testing.

>>>>

This paper proposes (1) a number of formal definitions of types of
nondeterminism and (2) a number of techniques, based on these definitions, for detecting and debugging
nondeterminism in property-based testing, particularly for library
code.  The methods also work for testing at the system level, but are
perhaps most effective for identifying the low-level sources of
nondeterminism in libraries.  This enables a number of useful results:
 the automated production of
nondeterminism-free regression tests even for libraries with some
sources of nondeterminism, the documentation of library
nondeterminism, and, most importantly, detection and effective debugging of
unexpected nondeterminism that is, itself, a bug.  We first
describe the types of nondeterminism we aim to detect, and the
extension of a widely used debugging algorithm to better handle
nondeterminism as a test property, then describe an implementation of
these ideas in a Python testing tool that is used on real-world
projects.  Finally, we use a set
of case studies of real-world Python libraries to demonstrate the
utility of our ideas and understand the cost of adding nondeterminism
detection to existing automated testing.
